\documentclass[10pt,twoside,twocolumn,fleqn]{article}

\input{macro}
\input{macros_tobias}

\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
% \usepackage{eucal}
\usepackage{graphicx}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage{geometry}
\usepackage{palatino}
% \usepackage{mlapa}
% \usepackage{hyperref}


% \renewcommand{\baselinestretch}{1.1}
\geometry{a4paper,headsep=7mm,hdivide={15mm,*,15mm},vdivide={20mm,*,15mm}}



%headsep=0mm,
\allowdisplaybreaks

\fancyhead[OL,ER]{\thetitle \textit{Tobias Lang and Marc Toussaint}---\today}
\fancyhead[C]{}
\fancyhead[OR,EL]{\thepage}
\fancyfoot{}
\pagestyle{fancy}



\definecolor{codecol}{rgb}{.9,.9,.9}
\usepackage{listings}
\lstset{ %
    language=C++,                % choose the language of the code
    basicstyle=\sf\footnotesize,       % the size of the fonts that are used
% for the code
    frame=none,                   % adds a frame around the code
    tabsize=4,                      % sets default tabsize to 2 spaces
    captionpos=b,                   % sets the caption-position to bottom
    texcl=true,
    mathescape=true,
    backgroundcolor=\color{codecol},
    escapechar=\%,
    columns=flexible,
    xleftmargin=1ex,
%    numbers=left, numberstyle=\footnotesize, stepnumber=1, numbersep=3ex
}




\mytitle{libARMANI -- Version 1.0\\Autonomous Robot Manipulation Simulator}
\myauthor{Tobias Lang and Marc Toussaint \\
\texttt{http://userpage.fu-berlin.de/tlang/armani/}}


\begin{document}
\maketitle

\begin{center}
\emph{When using this library, please cite \citet{lang-toussaint-10jair}.}
\end{center}

\tableofcontents



% ------------------------------------------------------------------------
% ------------------------------------------------------------------------
% ------------------------------------------------------------------------

\section{Introduction}

libARMANI is a C++-simulator for \textbf{Autonomous Robot MANIpulation
of objects}. It simulates a robot which manipulates cubes, balls and
boxes of different sizes scattered on a table. This provides a
physically realistic extension of one of the most popular scenarios in
Artificial Intelligence, namely of the good, old ``blocks world''.

libARMANI has been developed to serve as an easy-to-use test-bed for
\textbf{model-based relational reinforcement learning in stochastic
domains} \citep{lang-toussaint-10jair}: the actions of the simulated
robot are affected by noise; relational representations are appropriate
to learn and reason with the simulator objects.



% ------------------------------------------------------------------------
% ------------------------------------------------------------------------
% ------------------------------------------------------------------------

\section{Installation}

\subsection{Required Software}

\begin{itemize}
  \item Freeglut (tested with v2.4.0): probably already installed on your
system (Debian/Ubuntu: freeglut3-dev)
  \item QHull (tested with v2003.1): probably already installed on your system
(Debian/Ubuntu: libqhull-dev)
  \item ODE (tested with v0.11): probably needs to be installed by hand (make
sure to configure ODE with \texttt{--enable-double-precision}, which isn't
the default)
  \item SWIFT++ (tested with v1.2): comes with libARMANI in directory
\texttt{extern}
  \item ANN (tested with v1.1.1): probably already installed on your system
(Debian/Ubuntu: libann-dev)
\end{itemize}
You might need to compile the libraries, and libARMANI, with the flag -fPIC.



\subsection{Compilation}

First, please compile SWIFT++ in the directory \texttt{extern/}
and ensure that the correct pathes are set for Freeglut,
QHull, ODE and ANN in \texttt{make-generic}.

Then, please type \code{make} in the root-directory
\code{libARMANI/}. This should compile the library
\code{lib/libARMANI.so} and the demos in \code{test/}. If you have
problems getting this working, please contact me. If compilation
succeeded, you can check out the demos in \code{test/}.

You can clean everything via \texttt{libARMANI/make cleanAll}.



% ------------------------------------------------------------------------
% ------------------------------------------------------------------------
% ------------------------------------------------------------------------

\section{ORS Configuration Files}

Configuration files specify all objects and robot body parts used in the
simulator: from the robot's individual fingers over the table to the
manipulated balls. The configuration files reflect Marc Toussaint's
graph-based representation Open Robot Simulator Toolkit (ORS). ORS is
the backbone data representation of libARMANI. For more information,
please take a look at \texttt{http://userpage.fu-berlin.de/}
\texttt{mtoussai/source-code/}.

Example configuration files are
\texttt{test/relational\_armani\_basic/situation\_simple.ors} and
\texttt{test/relational\_armani\_basic/situation\_box.ors}.

The important part for your modifications is the list of manipulated
objects at the end of the file. Please note that these object
definitions need to stay at the end of the file.

\subsection{Cubes and balls}

How to specify cubes (blocks) and balls is best explained by means of an
example:

\begin{lstlisting}
body o1 { X=<t(-0.3 -0.7 0.8)> type=1 mass=.1
   size=[.03 .03 .03 .03] color=[0.4 0 .5] contact }
\end{lstlisting}

\texttt{o1} is the name of the object. Objects have to be named \texttt{oX}
where \texttt{X} is a number in ascending order.

\texttt{X=<t(-0.3 -0.7 0.8)>} specifies the position in (x/y/z)-coordinates.
(To get an idea of the coordinate system, play with the parameters.)

\texttt{type} specifies the type of the object which can either be a cube (0) or
a ball (1).

\texttt{size=[.03 .03 .03 .03]} specifies the size of the object. Use the
following sizes for objects:
\begin{itemize}
  \item Big cube: [.06 .06 .06 .06]
  \item Small cube: [.04 .04 .04 .04]
  \item Big ball: [.045 .045 .045 .045]
  \item Small ball: [.03 .03 .03 .03]
\end{itemize}

The mentioned requirements need to be met to be certain that all our
implementations work correctly. In principle, you are not obliged to
them, of course, in order to get the simulator run. But then you might
need to modify some of our provided implemented methods.


\subsection{Boxes}

\textbf{Boxes} are more complicated to specify: they consist of
combining an ORS-body with several ORS-shapes (the individual sides of
the box). I recommend to simply look at the example file
\texttt{test/relational\_armani\_basic/situation\_box.ors}.



% ------------------------------------------------------------------------
% ------------------------------------------------------------------------
% ------------------------------------------------------------------------

\section{User's Guide}

There are two important header files in \texttt{src/relational/}:
\begin{itemize}
\item \texttt{robotManipulationSimulator.h} declares the simulator. It
provides methods (i) to query the current state of the simulator and
(ii) to trigger actions of the simulated robot. The main actions are
grabbing objects and dropping them above other objects.

\item \texttt{robotManipulationInterface.h} provides a simple high-level
interface between the simulator and symbolic relational representations.
\end{itemize}



\subsection{Simulator Core}

At its heart, the simulator uses Marc Toussaint's graph-based
representation Open Robot Simulator Toolkit (ORS) and approximate
inference for control (AICO) \citep{toussaint:09-icml} to control the
robot. For more information, please take a look at 
\texttt{http://userpage.fu-berlin.de/}
\texttt{mtoussai/source-code/}, in
particular \texttt{libAICO}, where you'll find a guide explaining all
this in detail.

The simulator uses unsigned integer variables to identify objects.

The behavior of the robot is defined in
\texttt{robotManipulationSimulator}. For example, if we command the
robot to grab an object it first puts the inhand object on the table; if
it grabs an object below some other object, it will fail to grab this
object with 40\% probability. You may choose to change all these
behaviors according to your own ideas.

The concept of \textbf{noise} may be of particular interest for you to
implement your own ideas of stochastic worlds. Of course, the
simulator by itself is unnoisy and you have to specify explicitly
sources of noise. For example, we chose as a major source of noise the
calculation of a target location in a puton action (take a look at the
method \texttt{calcTargetPositionForDrop}). We deliberately add Gaussian
noise to the target location. You may want to define differents sources
and/or degrees of noise.




\subsection{Symbolic Relational Representations}
\label{programmerGuide:logic}

libARMANI provides the interface \texttt{robotManipulationInterface.h}
to symbolic relational representations. This interface uses relational
representations as implemented in \texttt{libPRADA}, a library for
model-based relational reinforcement learning. You can download this
library together with a guide from
\texttt{http://userpage.fu-berlin.de/tlang/prada/}.



\subsection{Demos}

The following demos are provided in \texttt{test/}:

\begin{itemize}
\item \texttt{relational\_armani\_basic/}: \\
This example program shows you how to directly control the simulator via
\texttt{robotManipulationSimulator.h}. Objects are managed by means
of their ids (which are unsigned integers), which are used to retrieve
state information and to trigger actions.

\item \texttt{relational\_armani\_symbolic/}: \\
This example shows you how to use an abstract relational
representation to control the robot via the interface in
\texttt{robotManipulationInterface.h}. This uses the extensive logic
machinery (see the various files in \texttt{src/relational/}) of our
experiments in \cite{lang-toussaint-10jair} which is implemented in
libPRADA. It provides many convenience functions (writing to files,
specifying symbols etc.). See Sec.~\ref{programmerGuide:logic}.
\end{itemize}






% ------------------------------------------------------------------------
% ------------------------------------------------------------------------
% ------------------------------------------------------------------------

\section{Trouble Shooting}

\subsection{ODE}

Make sure to configure ODE with \texttt{--enable-double-precision}, which isn't
the default.

In \texttt{ors\_ode.cpp} and \texttt{robotManipulationSimulator.cpp}, you
might want to change the following includes

\begin{lstlisting}
#  include <ode/../internal/objects.h>
#  include <ode/../internal/joints/joints.h>
#  include <ode/../internal/collision_kernel.h>
#  include <ode/../internal/collision_transform.h>
\end{lstlisting}

to:

\begin{lstlisting}
#  include <ode/../../ode/src/objects.h>
#  include <ode/../../ode/src/joints/joints.h>
#  include <ode/../../ode/src/collision_kernel.h>
#  include <ode/../../ode/src/collision_transform.h>
\end{lstlisting}

You might also want to change the LinuxLibs ODE library specification
appropriately to your needs.


\bibliographystyle{plainnat}
\bibliography{references}


\end{document}
